\vspace{-2cm}A generative model for urban firm networks

### 2020-04-22 main.tex
Update on Overleaf.
.

\title{An agent-based model for city networks based on interactions between firms}

@@

\author{J. Raimbault$^{1,2,3,\ast}$, N. Zdanowska$^{1,3}$ and E. Arcaute$^1$\\\medskip\small
$^{1}$ Centre for Advanced Spatial Analysis, UCL; $^{2}$ UPS CNRS 3611 ISC-PIF; $^{3}$ UMR CNRS 8504 G{\'e}ographie-cit{\'e}s\\
$^{\ast}$ \texttt{juste.raimbault@polytechnique.edu}
}

@@

\begin{abstract}
    This paper introduces a generative model for urban networks defined by interactions between firms to test several geographical properties emerging in such a network as metropolisation or internationalisation. The growth of the network is simulated by adding randomly new links defined by ownership between firms, with probabilities of occurring depending on the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past as well as the geographical and socio-cultural distance. The simulation on a synthetic system of cities unveil phase transitions when changing interaction distance, and non-trivial patterns in the model behavior. Calibration of the model on real data on company ownership linkages for Europe in 2018 is also undertaken. Future work will consist in observing thanks to this model the impact of economic crisis on such networks.

@@

The use of network models to understand processes underlying the emergence and dynamics of such urban networks has already been widely tackled in the literature. \cite{dai2016simulating} propose a generative urban network model at the regional scale, combining geographical factors with network topological properties. Air transportation networks can also be understood as urban networks, and \cite{zanin2013modelling} reviews several models for such networks. Trade network have been extensively studied from complex networks perspective, but are generally driven at the country-level \cite{garlaschelli2005structure} because of lack of city-level data. Input-output models, mostly used in regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}. In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for abstract flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2020indirect}.

%\citep{martinus2018global} combination of different types of proximity in inter-urban firm networks: economic, sociocultural, geopolitical - implies spatial and non-spatial 
%\citep{pan2017mapping} different types of linkages: e.g. services here, but not only. All sectors should be taken into account first.
%\cite{martinus2019brokerage} role of small countries
%\cite{dawley2019creating} regional policies for insertion in global network
%\cite{gluckler2016relational} positioning within networks

In this paper we choose to focus on urban networks of firms to question how can we understand and capture geographical and economic processes as agglomeration economies \cite{Glaeser2001} characteristic to spatial economy \cite{Fujitaetal2001} within a generative model. The originality of the present paper is to approach these issues within the system of cities framework \cite{berry1964cities}, where the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}. We examine the interactions of European cities within firm linkages defined by corporate ownership links. Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geographical structures \cite{2019arXiv191014652Z}.

@@

We then apply the model to a real system of cities by calibrating it on the European ownership network. Model setup is done following the real setup described above. The number of time steps $t_f$ is left as an additional parameter, which in a sense determines the granularity of the cumulative network generation process. The objective functions for calibration are the average mean squared error on logarithms of weights given by $\varepsilon_L = \frac{1}{N^2} \sum_{i,j} \left(\log w_{ij} - \log \hat{w}_{ij} \right)^2$ if $\hat{w}_{ij}$ are the simulated weights, and the logarithm of the mean squared error $\varepsilon_M = \log\left(\frac{1}{N^2} \sum_{i,j} \left(w_{ij} - \hat{w}_{ij}\right)^2 \right)$. \cite{raimbault2018indirect} showed that these two are complementary to calibrate urban systems models. Calibration is done using a genetic NSGA2 algorithm \cite{deb2002fast}, which is particularly suited for such a bi-objective calibration of a stochastic simulation model. We use a population of 200 and run the algorithm for 20,000 generations, on 1000 islands in parallel.

%%%%%%%%%%%%%%
\begin{figure}
	\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/pareto_colorgammaLinks.png}
    \end{center}
	\caption{Pareto front obtained for the bi-objective model calibration.\label{fig:calibration}}
\end{figure}
%%%%%%%%%%%%%%

We show in Fig.~\ref{fig:calibration} the calibration result, as a Pareto front of compromise points between the two objectives. These correspond to parameter values giving network trajectories closest to reality. The values for $\varepsilon_L$ are always lower than for statistical models, our model outperforming these in this predictive sense.
% add discussion param values
This furthermore shows in particular a role of the path-dependency parameter, which calibrated value is quite high (8 in average on the Pareto front), confirming the relevance of this complex generative model including path-dependency. These results altogether demonstrate the feasibility of model calibration on real data.

@@

This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMole platform and calibrated on real data on ownership linkages of firms for Europe. The simulation on a synthetic system of cities unveil phase transitions when changing interaction distance, and non-trivial patterns in the model behavior. This kind of model can have practical application in the future in terms of testing the effect of exogenous shocks in the socio-economic structure as the upcoming effect of Brexit on the European market.

@@

% Several scenarios are considered taking into consideration the current and future economic challenges, such as those given by Brexit. The model may be applied in future work to assess policy issues and recommendations.

% - other formulation of the combination of factors
%multi-objective optimisation depending on sectors, using Pareto fronts ? - needs empirical evidence; (iv) copula-based combination as done by \cite{2019arXiv190505106C} - copula parameters are free or estimated from real data
% The function used can be understood as a separable spatial interaction model. As a possible extension, Copula capture a dependency structure but must be parametrised by estimation on real data.

% - evolution of city sizes (co-evolution model)
% - role of path dependency
% - towards a model with firm agents? (multi-scale ABM)

\bibliographystyle{apalike}
%\bibliographystyle{unsrt}
\bibliography{biblio}

>>>>

\title{\vspace{-2cm}A generative model for urban firm networks}

@@

\author{J. Raimbault$^{1,2,3,\ast}$, N. Zdanowska$^{1,3}$\smallskip\\
$^{1}$ Centre for Advanced Spatial Analysis, UCL\\
$^{2}$ UPS CNRS 3611 ISC-PIF\\
$^{3}$ UMR CNRS 8504 G{\'e}ographie-cit{\'e}s\\
$^{\ast}$ \texttt{juste.raimbault@polytechnique.edu}
}

@@

\begin{abstract}
The emergence of entangled urban networks is a salient feature of globalisation processes. Understanding the drivers behind the growth of such networks, and in particular urban firm networks, is crucial for the economic resilience of urban systems. We introduce in this paper a generative network model for firm networks at the urban area level, which includes several complementary processes which are the economic size of urban areas at origin and destination, industrial sector proximity between firms, the strength of links from the past as well as the geographical and socio-cultural distance. An empirical network analysis on European firm ownership data confirms a role for each factor. We then simulate network growth for synthetic systems of cities, unveiling stylized facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range. We finally calibrate the model on the European network, outperforming statistical models and showing a strong role of path-dependency. Potential applications of the model include the study of mitigation policies for exogenous shocks such as economic crisis.

@@

The use of network models to understand processes underlying the emergence and dynamics of such urban networks has already been widely tackled in the literature. \cite{dai2016simulating} propose a generative urban network model at the regional scale, combining geographical factors with network topological properties. Air transportation networks can also be understood as urban networks, and \cite{zanin2013modelling} reviews several models for such networks. Trade network have been extensively studied from complex networks perspective, but are generally driven at the country-level \cite{garlaschelli2005structure} because of lack of city-level data. Input-output models, mostly used in regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}. In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for abstract flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2018indirect}.

In this paper we choose to focus on urban networks of firms to question how can we understand and capture geographical and economic processes as agglomeration economies \cite{Glaeser2001} characteristic to spatial economy \cite{Fujitaetal2001} within a generative model. The specific question we tackle is the complementarity of different processes to generate such urban networks. One originality of the present paper is to approach these issues within the system of cities framework \cite{berry1964cities}, where the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}. We examine the interactions of European cities within firm linkages defined by corporate ownership links. Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geographical structures \cite{2019arXiv191014652Z}.

@@

We then apply the model to a real system of cities by calibrating it on the European ownership network. Model setup is done following the real setup described above. The number of time steps $t_f$ is left as an additional parameter, which in a sense determines the granularity of the cumulative network generation process. The objective functions for calibration are the average mean squared error on logarithms of weights given by $\varepsilon_L = \frac{1}{N^2} \sum_{i,j} \left(\log w_{ij} - \log \hat{w}_{ij} \right)^2$ if $\hat{w}_{ij}$ are the simulated weights, and the logarithm of the mean squared error $\varepsilon_M = \log\left(\frac{1}{N^2} \sum_{i,j} \left(w_{ij} - \hat{w}_{ij}\right)^2 \right)$. \cite{raimbault2018indirect} showed that these two are complementary to calibrate urban systems models. Calibration is done using a genetic NSGA2 algorithm \cite{deb2002fast}, which is particularly suited for such a bi-objective calibration of a simulation model. Stochasticity is taken into account by adding the estimation confidence as an additional objective, and we filter the final population to have at least 20 samples. We use a population of 200 individuals and run the algorithm for 20,000 generations, on 1000 islands in parallel.

%%%%%%%%%%%%%%
\begin{figure}
	\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/pareto_colorgammaOrigin.png}
    \end{center}
	\caption{Pareto front obtained for the bi-objective model calibration for $\varepsilon_L$ and $\varepsilon_M$ objectives. $\varepsilon_L$ can be compared to the MSE of statistical models fitted above, and the generative model outperform these for several parametrizations (lowest MSE 5.33 for statistical models). Point size gives the number of stochastic samples and color the value of $\gamma_F$.\label{fig:calibration}}
\end{figure}
%%%%%%%%%%%%%%

We show in Fig.~\ref{fig:calibration} the calibration result, as a Pareto front of compromise points between the two objectives. As the mean squared error on logarithm is the same than the one for statistical models, it can be directly compared. The other objective would correspond to statistical models directly on variables: therefore, our model covers in a sense a set of intermediate models which make a compromise between the two objectives, and is therefore much more flexible. The values for $\varepsilon_L$ are for a large proportion of points lower than for statistical models, our model outperforming these in this predictive sense. As shown in Fig.~\ref{fig:calibration}, increasing values for $\gamma_F$ lead to better performance regarding $\varepsilon_L$: increasing the importance of the origin will provide a better fit on all order of magnitudes of link weights (while $\varepsilon_M$ favors mostly the largest links given the fat-tail distribution).

When considering optimal points with $\varepsilon_L < 5.0$, the adjusted parameters are: long range gravity interactions and socio-economic interactions given by $d_G = 1/\gamma_G = 5205 \pm 1775$ and $d_D = 1/ \gamma_D = 109 \pm 29$ (the maximal $g_{ij}$ is 11.8); a stronger influence of destination than origin as $\gamma_F = 1.45 \pm 0.09$ and $\gamma_T = 2.0 \pm 0.06$; a comparable influence of sectors as $\gamma_S = 1.75 \pm 0.17$; a fine time granularity as $t_f = 4960 \pm 109$; and a very strong influence of previous links as $\gamma_W = 7.99 \pm 3.77$. This last point furthermore shows in particular a role of the path-dependency parameter, confirming the relevance of this complex generative model including path-dependency. The interaction of this process with others implies an inversion of the relative importance of origin and destination (on the contrary to statistical models). The fact that calibrated interaction distances are large is consistent with long range economic interactions within Europe. Besides these stylized facts, these calibration results altogether demonstrate the feasibility of model application to real data.

@@

This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMole platform and calibrated on real data on ownership linkages of firms for Europe. The simulation on a synthetic system of cities provides an broad knowledge on model behavior in its parameter space, and its calibration suggests its potential application to policy issues. This kind of model can indeed have practical applications in the future in terms of testing the effect of exogenous shocks in the socio-economic structure as the upcoming effect of Brexit on the European market.

@@

\newpage

%\bibliographystyle{apalike}
\bibliographystyle{unsrt}
\bibliography{biblio}
